---
title: "DS-301-01-1_09_09_26"
format: html
editor: visual
---

## EDA for Comp Sci Plus One

## Loading Libraries

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(bayesrules)
library(ggpubr)
library(plotly)
library(report)
```

```{r}
df = read.csv("HT_2013-2025.csv")
```

## Plot 1

```{r}
#number of each sub offense
total <- sum(df$ACTUAL_COUNT, na.rm = TRUE)

df %>%
  group_by(OFFENSE_SUBCAT_NAME) %>%
  summarize(total = sum(ACTUAL_COUNT, na.rm = TRUE)) %>%
  ggplot(aes(x = OFFENSE_SUBCAT_NAME, y = total)) +
  geom_col(fill = "steelblue") +
  theme_minimal() +
  labs(
    title = "Total Count by Category",
    x = "Category",
    y = "Total Count"
  )
```

```{r}
summary <- df %>%
  group_by(DATA_YEAR) %>%
  summarise(
    distinct_agencies = n_distinct(ORI),
    total_incidents = sum(ACTUAL_COUNT, na.rm = TRUE)
  )

year <- summary %>%
  pivot_longer(
    cols = c(distinct_agencies, total_incidents),
    names_to = "Metric",
    values_to = "Value"
  )

ggplot(year, aes(x = DATA_YEAR, y = Value, color = Metric)) +
  geom_line(linewidth = 1) +
  geom_point() +
  scale_y_log10(labels = scales::comma) +
  labs(
    title = "Reporting Agencies vs. Total Incidents by Year",
    x = "Year",
    y = "Count (Log Scale)",
    color = "Metric"
  ) +
  theme_minimal()
```

```{r}
zeros <- df %>% filter(ACTUAL_COUNT == 0)
#unique agencies with 0 in actual count row
zeros %>%
  summarise(
    n_zero_rows = n(),
    n_agencies  = n_distinct(ORI)
  )

#agencies that have reported 0 in acutal count by year
df %>%
  group_by(DATA_YEAR) %>%
  summarise(
    n_agencies      = n_distinct(ORI),
    n_zero_rows     = sum(ACTUAL_COUNT == 0, na.rm = TRUE),
    pct_zero        = mean(ACTUAL_COUNT == 0, na.rm = TRUE),
    total_incidents = sum(ACTUAL_COUNT, na.rm = TRUE)
  ) %>%
  arrange(desc(pct_zero))

```

## Plot 2

```{r}
df %>%
  count(DATA_YEAR) %>%
  ggplot(aes(x = reorder(DATA_YEAR, -n), y = n)) +
  geom_col() +
  labs(x = "Year", y = "Count", title = "recorded incidents per year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
df |>
  group_by(DATA_YEAR) |>
  summarise(total = sum(ACTUAL_COUNT, na.rm = TRUE)) |>
  # factor() because DATA_YEAR is an integer: this gives one bar per year
  # instead of a continuous axis with half-year tick marks
  ggplot(aes(x = factor(DATA_YEAR), y = total)) +
  geom_col(fill = "steelblue") +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Year", y = "Total incidents",
       title = "Reported human trafficking incidents per year",
       caption = "Source: FBI UCR, 2013-2025") +
  theme_minimal()
```

## Plot 3

```{r}
#number of reports in each state
df %>%
  count(STATE_NAME) %>%
  ggplot(aes(x = reorder(STATE_NAME, -n), y = n)) +
  geom_col() +
  labs(x = "State Name", y = "Count", title = "Frequency of State") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| fig-width: 8
#| fig-height: 5

df %>%
  group_by(STATE_NAME) |> 
  summarise(total = sum(ACTUAL_COUNT, na.rm = TRUE)) |> 
  slice_max(total, n = 15) |> 
  ggplot(aes(x = total, y = fct_reorder(STATE_NAME, total))) +
  geom_col(fill = "steelblue") +
  scale_x_continuous(labels = scales::comma,
                     expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Total incidents", y = NULL,
       title = "Top 15 states by reported incidents, 2013-2025",
       subtitle = "These 15 states account for 63% of all reported incidents",
       caption = "Source: FBI UCR") +
  theme_minimal()
```

### Incidents vs. reporting agencies by state

```{r}
#| fig-width: 10
#| fig-height: 10

df %>%
  group_by(STATE_NAME) |> 
  summarise(Incidents = sum(ACTUAL_COUNT, na.rm = TRUE), # number of incidents
            `Reporting agencies` = n_distinct(ORI)) |> # number of reporting agencies
  pivot_longer(c(Incidents, `Reporting agencies`),
               names_to = "metric", values_to = "value") |> 
  mutate(STATE_NAME = fct_reorder2(STATE_NAME, metric, value)) |> 
  ggplot(aes(x = value, y = STATE_NAME)) +
  geom_col(fill = "steelblue") +
  facet_wrap(~ metric, scales = "free_x") +
  labs(x = NULL, y = NULL,
       title = "Incidents and reporting agencies by state, 2013-2025",
       subtitle = "States with more reporting agencies tend to report more incidents") +
  theme_minimal()
```

## Plot 4 - Subcategory

```{r}
# SUBCATEGORY -- stacked bar by year
df |> 
  group_by(DATA_YEAR, OFFENSE_SUBCAT_NAME) |> 
  summarise(offenses = sum(ACTUAL_COUNT), .groups = "drop") |> 
  ggplot(aes(x = factor(DATA_YEAR), y = offenses, fill = OFFENSE_SUBCAT_NAME)) +
  geom_col() +
  labs(title = "Offenses by year and subcategory",
       x = "Year", y = "Offenses", fill = "Subcategory") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Analysis:

```{r}
state_df <- df %>% 
  group_by(STATE_ABBR, OFFENSE_SUBCAT_NAME) %>%
summarise(COUNT = sum(ACTUAL_COUNT), .groups = 'keep')
model <- lm(COUNT ~ OFFENSE_SUBCAT_NAME + STATE_ABBR, data = state_df)
summary(model)
```

```{r}
#| fig-width: 10
#| fig-height: 6
ggplot(state_df, aes(x = STATE_ABBR, y = COUNT, color = OFFENSE_SUBCAT_NAME)) +
  geom_jitter(width = .1, alpha = .5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_smooth(method = "lm") +
  stat_cor()
```

```{r}
year_df <- df %>% 
  group_by(DATA_YEAR, OFFENSE_SUBCAT_NAME) %>%
  summarise(COUNT = sum(ACTUAL_COUNT), .groups = 'keep')
model <- lm(COUNT ~ OFFENSE_SUBCAT_NAME + DATA_YEAR, data = year_df)
summary(model)

```

```{r}
#| fig-width: 10
#| fig-height: 6
ggplot(year_df, aes(x = DATA_YEAR, y = COUNT, color = OFFENSE_SUBCAT_NAME)) +
  geom_jitter(width = .1, alpha = .5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_smooth(method = "lm") +
  stat_cor()
```

## Analysis Caption:

This graph is promising, as it shows the relationship between the offense subcategory types over time. We can see a linear growth in both offense types (which overall is not good) however the red line, which represents commercial sex acts has a significantly higher slope than the involuntary servitude offense type. Spending our next project stages honing in on why there is such a big difference in slope is something that our group will definitely look into. Each point on the graph represents the total number of incidences that occurred under that offense subcategory type in that year. This means there are only 2 points every 6 months that are being shown in this graph. I think that this graph could definitely be useful in building our story because it shows the concerning rate of change in both offense subcategory types as time goes on. There is a good chance this visualization will remain a staple of our project and narrative.

```{r}
# 1. Prepare sample data
# Already using df #

levels(df$OFFENSE_SUBCAT_NAME)
class(df$OFFENSE_SUBCAT_NAME)

df$OFFENSE_SUBCAT_NAME_F <- as.factor(df$OFFENSE_SUBCAT_NAME)
class(df$OFFENSE_SUBCAT_NAME_F)
levels(df$OFFENSE_SUBCAT_NAME_F)

# 2. Fit the linear model
model <- lm(ACTUAL_COUNT ~ OFFENSE_SUBCAT_NAME, data = df)

# 3. View the regression coefficients and statistical summary
summary(model)

```

```{r}
ggplot(df, aes(x = OFFENSE_SUBCAT_NAME, y = ACTUAL_COUNT)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor()

ggplot(df, aes(x = OFFENSE_SUBCAT_NAME, y = ACTUAL_COUNT)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.5) +
  geom_smooth(method = "lm") +
  stat_cor()
```

## Analysis Caption:

This scatter plot shows the distribution of how many incidences are occurring per row of data inside the data set. This visualization also splits filters the data by offense subcategory name. Most of the rows include incidences closer to 10-20 as shown by the density of the points towards the bottom of the graph. This was good practice to work with the data, although at the moment I am unsure of how I would use this to build the narrative for our presentation. I am leaning towards not bringing this visualization forward into future stages of the project process but it is better to have an abundance of visualizations rather than not enough.
