---
title: "DS-301-01-1_09_09_26"
format: html
editor: visual
---

## EDA for Comp Sci Plus One

## Loading Libraries

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(bayesrules)
library(ggpubr)
library(plotly)
library(report)
```

```{r}
df = read.csv("HT_2013-2025.csv")
```

## Plot 1

```{r}
#number of each sub offense
total <- sum(df$ACTUAL_COUNT, na.rm = TRUE)

df %>%
  group_by(OFFENSE_SUBCAT_NAME) %>%
  summarize(total = sum(ACTUAL_COUNT, na.rm = TRUE)) %>%
  ggplot(aes(x = OFFENSE_SUBCAT_NAME, y = total)) +
  geom_col(fill = "steelblue") +
  theme_minimal() +
  labs(
    title = "Total Count by Category",
    x = "Category",
    y = "Total Count"
  )
```

```{r}
summary <- df %>%
  filter(ACTUAL_COUNT != 0) %>% 
  group_by(DATA_YEAR) %>%
  summarise(
    distinct_agencies = n_distinct(ORI),
    total_incidents = sum(ACTUAL_COUNT, na.rm = TRUE)
  )

year <- summary %>%
  pivot_longer(
    cols = c(distinct_agencies, total_incidents),
    names_to = "Metric",
    values_to = "Value"
  )

ggplot(year, aes(x = DATA_YEAR, y = Value, color = Metric)) +
  geom_line(linewidth = 1) +
  geom_point() +
  scale_y_log10(labels = scales::comma) +
  labs(
    title = "Reporting Agencies vs. Total Incidents by Year",
    x = "Year",
    y = "Count",
    color = "Metric"
  ) +
  theme_minimal()
```

```{r}
zeros <- df %>% filter(ACTUAL_COUNT == 0)
#unique agencies with 0 in actual count row
zeros %>%
  summarise(
    n_zero_rows = n(),
    n_agencies  = n_distinct(ORI)
  )

#agencies that have reported 0 in acutal count by year
df %>%
  group_by(DATA_YEAR) %>%
  summarise(
    n_agencies      = n_distinct(ORI),
    n_zero_rows     = sum(ACTUAL_COUNT == 0, na.rm = TRUE),
    pct_zero        = mean(ACTUAL_COUNT == 0, na.rm = TRUE),
    total_incidents = sum(ACTUAL_COUNT, na.rm = TRUE)
  ) %>%
  arrange(desc(pct_zero))

```

## Plot 2

```{r}
df %>%
  count(DATA_YEAR) %>%
  ggplot(aes(x = reorder(DATA_YEAR, -n), y = n)) +
  geom_col() +
  labs(x = "Year", y = "Count", title = "recorded incidents per year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Plot 3

```{r}
#number of reports in each state
# df %>%
#   count(STATE_ABBR) %>%
#   ggplot(aes(x = reorder(STATE_ABBR, -n), y = n)) +
#   geom_col() +
#   labs(x = "State", y = "Count", title = "Frequency of State") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1))

start_year <- 2016   # a researcher decision (see below)

window_df <- df |>
  filter(DATA_YEAR >= start_year)

n_years <- n_distinct(window_df$DATA_YEAR)

state_counts <- panel_df |>
  group_by(STATE_NAME) |>
  summarise(
    n_rows     = n(),                      # rows (agency-year-offense records)
    n_agencies = n_distinct(ORI),          # agencies in the balanced panel
    total_count = sum(ACTUAL_COUNT),       # offenses contributed to the totals
    .groups = "drop"
  ) |>
  arrange(desc(n_rows))

state_counts

# 2. Yearly totals by subcategory
yearly <- panel_df |>
  group_by(DATA_YEAR, STATE_NAME, OFFENSE_SUBCAT_NAME) |>
  summarise(total = sum(ACTUAL_COUNT), .groups = "drop") |>
  filter(total > 0)   # log(0) is -Inf

yearly$yr <- yearly$DATA_YEAR - start_year

lm(log(total) ~ yr * OFFENSE_SUBCAT_NAME + STATE_NAME, data = yearly)

#model tells us that texas is the main contributer to the data with 9 of there agencies participating in the reports which is vastly more than the other states with the next closest at 3
```

```{r}
library(dplyr)
library(ggplot2)

model <- lm(log(total) ~ yr * OFFENSE_SUBCAT_NAME + STATE_NAME, data = yearly)

cf <- coef(model)
state_eff <- cf[grepl("^STATE_NAME", names(cf))]

plot_df <- tibble(
  state = sub("^STATE_NAME", "", names(state_eff)),
  mult  = exp(state_eff)
) |>
  add_row(state = paste0(levels(factor(yearly$STATE_NAME))[1], " (ref)"), mult = 1)

ggplot(plot_df, aes(x = mult, y = reorder(state, mult))) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  geom_segment(aes(x = 1, xend = mult, yend = state)) +
  geom_point(size = 3) +
  scale_x_log10(breaks = c(0.25, 0.5, 1, 2, 4, 8, 16)) +
  labs(x = "Amount of reports reletive to log scale", y = NULL,
       title = "State differences") +
  theme_minimal()
```

```{r}
panel_df |>
  group_by(STATE_NAME) |>
  summarise(total = sum(ACTUAL_COUNT)) |>
  mutate(share = total / sum(total)) |>
  ggplot(aes(x = share, y = reorder(STATE_NAME, share))) +
  geom_col() +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Share of all offenses", y = NULL) +
  theme_minimal()
```

```{r}
by_year <- function(data) {
  data |>
    filter(ACTUAL_COUNT != 0) |>     # drops zero rows (and NAs)
    group_by(DATA_YEAR) |>
    summarise(
      n_agencies = n_distinct(ORI),
      total      = sum(ACTUAL_COUNT),
      .groups = "drop"
    )
}

all_states <- by_year(window_df)
no_texas   <- by_year(window_df |> filter(STATE_NAME != "Texas"))

cor.test(all_states$n_agencies, all_states$total)
cor.test(no_texas$n_agencies,   no_texas$total)

# state_year <- window_df |>
#   group_by(STATE_NAME, DATA_YEAR) |>
#   summarise(n_agencies = n_distinct(ORI), total = sum(ACTUAL_COUNT), .groups = "drop")
# 
# state_year |>
#   group_by(STATE_NAME) |>
#   summarise(r = cor(n_agencies, total), n_years = n())
```

## Plot 4 - Subcategory

```{r}
# SUBCATEGORY -- stacked bar by year
df |> 
  group_by(DATA_YEAR, OFFENSE_SUBCAT_NAME) |> 
  summarise(offenses = sum(ACTUAL_COUNT), .groups = "drop") |> 
  ggplot(aes(x = factor(DATA_YEAR), y = offenses, fill = OFFENSE_SUBCAT_NAME)) +
  geom_col() +
  labs(title = "Offenses by year and subcategory",
       x = "Year", y = "Offenses", fill = "Subcategory") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Analysis:

```{r}
state_df <- df %>% 
  group_by(STATE_ABBR, OFFENSE_SUBCAT_NAME) %>%
summarise(COUNT = sum(ACTUAL_COUNT), .groups = 'keep')
model <- lm(COUNT ~ OFFENSE_SUBCAT_NAME + STATE_ABBR, data = state_df)
summary(model)
```

```{r}
#| fig-width: 10
#| fig-height: 6
ggplot(state_df, aes(x = STATE_ABBR, y = COUNT, color = OFFENSE_SUBCAT_NAME)) +
  geom_jitter(width = .1, alpha = .5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_smooth(method = "lm") +
  stat_cor()
```

```{r}
year_df <- df %>% 
  group_by(DATA_YEAR, OFFENSE_SUBCAT_NAME) %>%
  summarise(COUNT = sum(ACTUAL_COUNT), .groups = 'keep')
model <- lm(COUNT ~ OFFENSE_SUBCAT_NAME + DATA_YEAR, data = year_df)
summary(model)

```

```{r}
#| fig-width: 10
#| fig-height: 6
ggplot(year_df, aes(x = DATA_YEAR, y = COUNT, color = OFFENSE_SUBCAT_NAME)) +
  geom_jitter(width = .1, alpha = .5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_smooth(method = "lm") +
  stat_cor()
```

## Analysis Caption:

This graph is promising, as it shows the relationship between the offense subcategory types over time. We can see a linear growth in both offense types (which overall is not good) however the red line, which represents commercial sex acts has a significantly higher slope than the involuntary servitude offense type. Spending our next project stages honing in on why there is such a big difference in slope is something that our group will definitely look into. Each point on the graph represents the total number of incidences that occurred under that offense subcategory type in that year. This means there are only 2 points every 6 months that are being shown in this graph. I think that this graph could definitely be useful in building our story because it shows the concerning rate of change in both offense subcategory types as time goes on. There is a good chance this visualization will remain a staple of our project and narrative.

```{r}
# 1. Prepare sample data
# Already using df #

levels(df$OFFENSE_SUBCAT_NAME)
class(df$OFFENSE_SUBCAT_NAME)

df$OFFENSE_SUBCAT_NAME_F <- as.factor(df$OFFENSE_SUBCAT_NAME)
class(df$OFFENSE_SUBCAT_NAME_F)
levels(df$OFFENSE_SUBCAT_NAME_F)

# 2. Fit the linear model
model <- lm(ACTUAL_COUNT ~ OFFENSE_SUBCAT_NAME, data = df)

# 3. View the regression coefficients and statistical summary
summary(model)

```

```{r}
ggplot(df, aes(x = OFFENSE_SUBCAT_NAME, y = ACTUAL_COUNT)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor()

ggplot(df, aes(x = OFFENSE_SUBCAT_NAME, y = ACTUAL_COUNT)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.5) +
  geom_smooth(method = "lm") +
  stat_cor()
```

## Analysis Caption:

This scatter plot shows the distribution of how many incidences are occurring per row of data inside the data set. This visualization also splits filters the data by offense subcategory name. Most of the rows include incidences closer to 10-20 as shown by the density of the points towards the bottom of the graph. This was good practice to work with the data, although at the moment I am unsure of how I would use this to build the narrative for our presentation. I am leaning towards not bringing this visualization forward into future stages of the project process but it is better to have an abundance of visualizations rather than not enough.
