---
title: "EDA — [your project name]"
author: "[group members]"
date: today
format:
  html:
    toc: true
    embed-resources: true
    fig-width: 7
    fig-height: 4.5
editor: visual
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)

# Load your data here. 

# dat <- read_csv("your_file.csv")
```

## What we have

Replace this with: how many rows, how many columns, and whether the row count matches what the documentation says. If it doesn't match, say so. That's a finding to observe, not a problem to hide.

```{r}
#| label: dimensions

# dim(dat)
# glimpse(dat)
```

------------------------------------------------------------------------

## Example

*This section is a model. Delete it before you submit.*

```{r}
#| label: fig-example
#| echo: true

ggplot(mpg, aes(x = hwy)) +
  geom_histogram(binwidth = 2) +
  labs(x = "Highway MPG", y = "Number of vehicles") +
  theme_minimal()
```

**Figure 0.** Distribution of highway fuel economy across 234 vehicle models, 1999–2008.

The distribution is right-skewed with a long upper tail, so the median (24 mpg) describes a typical vehicle better than the mean (23.4 mpg pulled toward the outliers). There's a small cluster above 40 mpg that is worth checking . Those may be a distinct vehicle class rather than extreme values of the same group. This describes only the models in this dataset, which is not a random sample of vehicles on the road.

> Notice what the interpretation does: names the shape, justifies the summary choice, flags something to follow up on, and scopes the claim. Yours should do the same things.

------------------------------------------------------------------------

## Figure 1: (list variables)

```{r}
#| label: fig-one

```

**Figure 1.** *Caption here: what the reader is looking at, including n and any exclusions.*

*Interpretation here: 2–3 sentences. What do you see? What does it NOT show?*

## Figure 2: (list variables)

```{r}
#| label: fig-two

```

**Figure 2.** *Caption.*

*Interpretation.*

## Figure 3: (list variables)

```{r}
#| label: fig-three

```

**Figure 3.** *Caption.*

*Interpretation.*

## One column, one instrument

Pick one column and answer in prose:

- **Column:**
- **What produced this number** (what instrument, question, sensor, or process):
- **What construct it was meant to capture:**
- **What you'd check to be sure** (usually the codebook...name the page if you found it):

# Optional: Starting analysis

Only if you're ready. Fill these in *before* running anything.

**What we expect:** *We expect \_\_\_ to be higher/lower in \_\_\_ than , because \_\_* .

**What would change our mind:** *If we see \_\_\_, our expectation was wrong, and \_\_\_.*

```{r}
#| label: analysis



```

**What we found** (including if it was nothing! A predicted null is still soemthing to talk about):
